---
title: Demonstration of matching route networks
format: gfm
---


```{r}
library(stplanr)
library(sf)
rnet_y = route_network_small["flow"]
rnet_y$id = 1:nrow(rnet_y)
# The target object
rnet_x = rnet_subset(osm_net_example[1], rnet_y)
# Get start and end points for each linestring in rnet_x:
rnet_x_s = lwgeom::st_startpoint(rnet_x)
rnet_x_e = lwgeom::st_endpoint(rnet_x)
plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
plot(rnet_y["flow"], add = TRUE, lwd = 5)
plot(rnet_x_s, add = TRUE, pch = 3)
plot(rnet_x_e, add = TRUE, pch = 3)
```
Here are the examples of an intersection road.

```{r}
rnet_intersection_complex = sf::read_sf("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_complex.geojson")
rnet_intersection_simple = sf::read_sf("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_simple.geojson")
rnet_intersection_simple$index = seq_along(rnet_intersection_simple$geometry)

rnet_intersection_simple = rnet_intersection_simple[, c("index", setdiff(names(rnet_intersection_simple), "index"))]


rnet_intersection_simple <- rnet_intersection_simple[c("index", "geometry")]

plot(rnet_intersection_complex$geometry, lwd = 20, col = "blue")
plot(rnet_intersection_simple$geometry, col = "red", add = TRUE, lwd = 5)
```
This is where the rnet_merge function is hard to handle at the intersection road.

```{r}
rnet_xp = sf::st_transform(rnet_intersection_simple , "EPSG:27700")
rnet_yp = sf::st_transform(rnet_intersection_complex, "EPSG:27700")
names(rnet_xp)
# Extract column names from the rnet_yp
name_list = names(rnet_yp)

# Initialize an empty list
funs = list()

# Loop through each name and assign it a function based on specific conditions
for (name in name_list) {
  if (name == "geometry") {
    next  # Skip the current iteration
  } else if (name %in% c("gradient", "quietness")) {
    funs[[name]] = mean
  } else {
    funs[[name]] = sum
  }
}


dist = 10
angle = 5
rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist, funs = funs, max_angle_diff = angle, segment_lentgh = 10)  

library(tmap)
tmap_mode("view")

brks = c(0, 50, 100, 200, 500,1000,2000)
m1 = tm_shape(rnet_intersection_complex) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
m2 = tm_shape(rnet_merged) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
tmap_arrange(m1, m2, nrow = 1, sync = TRUE)

```


The stplanr solution was as follows:

```{r}
rnet_y$quietness <- rnorm(nrow(rnet_y))
funs <- list(flow = sum, quietness = mean)
rnet_merged <- rnet_merge(rnet_x[1], rnet_y[c("flow", "quietness")],
  dist = 9, segment_length = 20, funs = funs
)
plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_merged["flow"], add = TRUE, lwd = 2)
```

Note that this leaves gaps in the network.

Although this gap issue can be fixed by using a smaller segment_length, e.g. segment_length = 5, this is not a computationally efficient way to solve the problem.

```{r}
rnet_y$quietness <- rnorm(nrow(rnet_y))
funs <- list(flow = sum, quietness = mean)
rnet_merged <- rnet_merge(rnet_x[1], rnet_y[c("flow", "quietness")],
  dist = 9, segment_length = 5, funs = funs
)
plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_merged["flow"], add = TRUE, lwd = 2)
```

Also, the join syntax is a bit clunky.

## With `{rsgeo}`

Let's try with the `{rsgeo}` implementation:

```{r}
library(rsgeo)
rnet_x = rnet_x |>
  sf::st_transform("EPSG:27700")
rnet_y = rnet_y |>
  sf::st_transform("EPSG:27700")
rnet_y$value = rnet_y$flow
x <- as_rsgeo(rnet_x)
y <- as_rsgeo(rnet_y)
res <- rsgeo:::rnet_merge(x, y, 20, 0.2)  
excluded <- setdiff(1:length(x), res$from)
from <- rnet_x[res$from,]
to <- lapply(res$to, \(.i) rnet_y[.i,])
to_mean_value <- sapply(res$to, \(.i) mean(rnet_y$value[.i]))
rnet_matched_rsgeo = dplyr::bind_cols(from, value = to_mean_value)
plot(rnet_matched_rsgeo["value"], lwd = 3)
```

## With `{geos}`

Let's try doing the same thing but with `geos`.
First we'll reproject the data.

```{r}
rnet_x_projected = st_transform(rnet_x, "EPSG:27700")
rnet_y_projected = st_transform(rnet_y, "EPSG:27700")
```

Then we'll convert to `geos` objects.

```{r}
rnet_x_geos = geos::as_geos_geometry(rnet_x_projected)
rnet_y_geos = geos::as_geos_geometry(rnet_y_projected)
```

We'll create a buffer around `rnet_x`:

```{r}
params = geos::geos_buffer_params(end_cap_style = "flat")
rnet_x_buffer = geos::geos_buffer(rnet_x_geos, 9, params = params)
plot(rnet_x_buffer)
plot(rnet_y_geos, add = TRUE, col = "red", lwd = 2)
```

Now let's 'chop' the source geometry into segments that fit within the buffer:

```{r}
# Convert to linestring:
rnet_xbl = geos::geos_boundary(rnet_x_buffer)
rnet_xblb = geos::geos_buffer(rnet_xbl, 3, params = params)
rnet_xlbc = geos::geos_make_collection(rnet_xblb)
rnet_xlbcu = geos::geos_unary_union(rnet_xlbc)
rnet_y_remove = geos::geos_intersection(
  rnet_y_geos,
  rnet_xlbcu
)
plot(rnet_xlbcu, col = "lightgrey")
```

```{r}
plot(rnet_xlbcu, col = "lightgrey")
plot(rnet_y_remove, add = TRUE, col = "red", lwd = 2)
```

The red bits are the parts of the source geometry `rnet_y` that we *don't* want.
Let's get the bits that we *do* want:

```{r}
rnet_y_chopped = geos::geos_difference(
  rnet_y_geos,
  rnet_xlbcu
)
plot(rnet_xlbcu)
plot(rnet_x_buffer, add = TRUE, col = "lightgrey", border = NA)
plot(rnet_y_chopped, add = TRUE, col = "red", lwd = 2)
```

<!-- For every 'chopped' linestring there is at least one matching linestring in `rnet_y`.
Let's find them as follows: -->

```{r}
#| echo: false
#| eval: false
rnet_ycj = geos::geos_inner_join_keys(
  rnet_y_chopped,
  rnet_y_geos
)
plot(rnet_ycj)
```


<!-- That shows that we're not quite ready to join the data: we need to exclude the red bits that are outside the original buffer. -->

```{r}
#| echo: false
#| eval: false
multilinestring = geos::geos_unary_union(
    geos::geos_make_collection(c(
      geos::geos_make_linestring(1:3, 1:3),
      geos::geos_make_linestring(11:13, 11:13)
    ))
)
# This does not generate a linestring:
geos::geos_unnest(multilinestring)
# # Convert back to linestring with ???:
# multilinestring_sf = sf::st_as_sf(multilinestring)
# sf::st_cast(multilinestring_sf, "LINESTRING")
geos::geos_unnest(multilinestring, keep_multi = FALSE)
```

```{r}
rnet_ycl = geos::geos_unnest(rnet_y_chopped, keep_multi = FALSE)
rnet_ycj = geos::geos_inner_join_keys(
  rnet_ycl,
  rnet_y_geos
)
plot(rnet_ycj)
```

We can also join `rnet_y_chopped` and `rnet_ycl` to `rnet_x_buffer` to get the buffer geometry:

```{r}
rnet_ycj = geos::geos_inner_join_keys(
  rnet_y_chopped,
  rnet_x_buffer
)
nrow(rnet_ycj)
length(rnet_y_chopped)
length(unique(rnet_ycj$x))
length(unique(rnet_ycj$y))
plot(rnet_ycj)
rnet_y


rnet_yclj = geos::geos_inner_join_keys(
  rnet_ycl,
  rnet_x_buffer
)
nrow(rnet_yclj)
length(rnet_ycl)
plot(rnet_yclj)
rnet_y
```

```{r}
rnet_ycj2 = geos::geos_inner_join_keys(
  rnet_y_geos,
  rnet_x_buffer
)
waldo::compare(rnet_ycj, rnet_ycj2)
```

```{r}
```

Now let's join the dataframe versions:

```{r}
rnet_y_df = sf::st_drop_geometry(rnet_y)
rnet_y_df_expanded = rnet_y_df[rnet_ycj$x, ]
rnet_x_df = sf::st_drop_geometry(rnet_x)
rnet_x_df_expanded = rnet_x_df[rnet_ycj$y, ]
rnet_y_df_expanded$osm_id = rnet_x_df_expanded$osm_id
nrow(rnet_y_df_expanded)
nrow(rnet_x)
rnet_x_agg = rnet_y_df_expanded |>
  dplyr::group_by(osm_id) |>
  dplyr::summarise(
    flow = sum(flow),
    quietness = mean(quietness)
  )
rnet_x_joined = dplyr::left_join(
    rnet_x,
    rnet_x_agg,
)
plot(rnet_x_joined["flow"])
```

Let's compare the old and new joined flows:

```{r}
# place plots side by side
par(mfrow = c(1, 2))
plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_merged["flow"], add = TRUE, lwd = 2)

plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_x_joined["flow"], add = TRUE, lwd = 2)
par(mfrow = c(1, 1))
```

To package this up we used the following commands:

```{r}
#| eval: false
usethis::use_description()
usethis::use_package("geos")
```

We created a function that takes `rnet_x` and `rnet_y` and returns a dataframe with the joined data:

```{r}
rnet_x = rnet_x_projected
rnet_y = rnet_y_projected
distance = 9
dist_chop = 0.1
rnet_join = function(
    rnet_x,
    rnet_y,
    distance = 9,
    dist_chop = 0.1
    ) {
  # Check CRS is projected:
  stopifnot(
    sf::st_is_longlat(rnet_x) == FALSE,
    sf::st_is_longlat(rnet_y) == FALSE
  )
  rnet_x_geos = geos::as_geos_geometry(rnet_x)
  rnet_y_geos = geos::as_geos_geometry(rnet_y)
  params = geos::geos_buffer_params(end_cap_style = "flat")
  rnet_x_buffer = geos::geos_buffer(rnet_x_geos, distance, params = params)
  rnet_xbl = geos::geos_boundary(rnet_x_buffer)
  rnet_xblb = geos::geos_buffer(rnet_xbl, dist_chop, params = params)
  rnet_xlbc = geos::geos_make_collection(rnet_xblb)
  rnet_xlbcu = geos::geos_unary_union(rnet_xlbc)
  rnet_y_chopped = geos::geos_difference(
    rnet_y_geos,
    rnet_xlbcu
  )
  rnet_ycj = geos::geos_inner_join_keys(
    rnet_y_chopped,
    rnet_x_buffer
  )
  rnet_y_df = sf::st_drop_geometry(rnet_y)
  rnet_y_df_expanded = rnet_y_df[rnet_ycj$x, ]
  rnet_y_df_expanded$id_x = rnet_x_df[[1]][rnet_ycj$y]
  # Replace 'id_x' with the name of the first column in rnet_x_df:
  nx = which(names(rnet_y_df_expanded) == "id_x")
  new_name = names(rnet_x_df)[1]
  names(rnet_y_df_expanded)[nx] = new_name
  rnet_y_df_expanded
}
# Test it:
res = rnet_join(rnet_x, rnet_y)
res
res |>
  dplyr::group_by(id) |>
  dplyr::summarise(
    osm_ids = paste(osm_id, collapse = ", ")
  )

```

```{r}
#| eval: false
#| echo: false
# Calculate angle:
geos::
```

# Benchmark

Let's test it on a larger dataset:

```{r}
rnet_x = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_x_ed.geojson")
rnet_y = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_y_ed.geojson")
```

```{r}
plot(rnet_y["value"], lwd = 5)
```

```{r}
plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
plot(rnet_y["value"], add = TRUE, lwd = 5)
plot(rnet_x_s, add = TRUE, pch = 3)
plot(rnet_x_e, add = TRUE, pch = 3)
```

Let's compare the `stplanr` implementation with the new implementation:

```{r}
# Stplanr implementation:
system.time({
rnet_merged = rnet_merge(rnet_x, rnet_y["value"], dist = 9, segment_length = 20, funs = list(value = sum))
})
plot(rnet_merged["value"], lwd = 3)
```

```{r}
devtools::load_all()
rnet_xp = rnet_x |>
  sf::st_transform("EPSG:27700")
rnet_yp = rnet_y |>
  sf::st_transform("EPSG:27700") 
system.time({
rnet_matched = rnet_match(rnet_xp, rnet_yp, distance = 9, dist_chop = 1)
rnet_matched_agg = rnet_matched |>
  dplyr::group_by(identifier) |>
  dplyr::summarise(value = sum(value))
rnet_joined = rnet_x |>
  dplyr::left_join(rnet_matched_agg)
})
summary(rnet_joined)
# rnet_joined$value
plot(rnet_joined["value"], lwd = 3)
```


# With rsgeo


```{r}
remotes::install_github("josiahparry/rsgeo", ref = "rent")

rnet_x = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_x_ed.geojson") |> 
  sf::st_transform(27700)
rnet_y = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_y_ed.geojson") |> 
  sf::st_transform(27700)

plot(rnet_x)
library(rsgeo)


x <- as_rsgeo(rnet_x)
y <- as_rsgeo(rnet_y)


# 3rd arg is distance 4th is slope tolerance
# distance is in units of the input data
system.time({
res <- rsgeo:::rnet_merge(x, y, 20, 0.2)  

excluded <- setdiff(1:length(x), res$from)

from <- rnet_x[res$from,]
to <- lapply(res$to, \(.i) rnet_y[.i,])
to_mean_value <- sapply(res$to, \(.i) mean(rnet_y$value[.i]))

rnet_matched_rsgeo = dplyr::bind_cols(from, value = to_mean_value)
})

plot(rnet_matched_rsgeo["value"], lwd = 3)

```

```{r}
#| eval: false
#| echo: false
# to be joined on
plot(x[res$from], lwd = 2)
# joined onto x from y
plot(y[unique(unlist(res$to))], col = "red", lty = 3, add = TRUE, lwd = 2)
# not used
plot(x[excluded], col = "blue", lty = 3, add = TRUE)
```

Let's try matching the data:

```{r}

```